---
output:
  pdf_document: default
  html_document: default
---

```{r}
rm(list=ls())
library(tidyverse)
library(ggthemes)
data <- read_csv("cytof_one_experiment.csv")
```

```{r}
## 1.)
data <- data %>%
          pivot_longer(names(data), names_to="Protein_Identity", values_to="Protein_Level")
data
nrow(data) == 1750000
```

```{r}
## 2.)
median_and_mad <- data %>%
                    group_by(Protein_Identity) %>%
                    summarise(Median=median(Protein_Level), MAD=mad(Protein_Level))
median_and_mad
                          
```

```{r}
## 3.)
ggplot(median_and_mad, aes(x=MAD, y=Median)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, col="red") +
  xlab("Median Absolute Deviation") +
  ylab("Median") +
  ggtitle("Spread-Location Plot") +
  theme_few()

cor(median_and_mad$MAD, median_and_mad$Median)
```

The spread-location plot is a null plot; there is no clear association between the median absolute deviation and the median. This is noteworthy because the median absolute deviation is just a robust version of the median. The lack of a pattern indicates that the median is unstable in this dataset, and the median absolute deviation should be used as a measure of center. 

```{r}
## 4.) 
library(dcldata)
data(example_gymnastics_2)
example_gymnastics_2 <- example_gymnastics_2 %>%
  pivot_longer(names(example_gymnastics_2)[-1], names_to="event", values_to="score")
example_gymnastics_2 <- example_gymnastics_2 %>% 
  separate(event, into = c("event", "year"))
example_gymnastics_2 <- example_gymnastics_2 %>%
  arrange(desc(score))
example_gymnastics_2
```
